---
title: "Week 3 Challenge"
author: "Business Science, student Elena Thomson"
date:   "21/01/2023"
output: 
    html_document:
        toc: TRUE
        theme: flatly
        highlight: tango
        code_folding: show
        df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE
    )
```

# Challenge Summary

This is a short challenge to begin applying what you are learning to the problem at hand. You will go through a series of questions related to the course project goals: 

1. Coming up with a new product idea, and 

2. Segmenting the customer-base

# Objectives

1. Apply `lubridate` and `stringr` functions to answer questions related to the course projects. 

2. Gain exposure to `rmarkdown`.

# Data

To read the data, make sure that the paths point to the appropriate data sets. Saving the file in the "challenges folder" should enable the paths to be detected correctly. 

```{r, message=FALSE, warning=FALSE}
# Load libraries
library(tidyverse)
library(lubridate)
```

```{r}
# Read bike orderlines data
path_bike_orderlines <- "../00_data/bike_sales/data_wrangled/bike_orderlines.rds"
bike_orderlines_tbl <- read_rds(path_bike_orderlines) %>%
    
    # Fix typos found in Feature Engineering
    mutate(model = case_when(
        model == "CAAD Disc Ultegra" ~ "CAAD12 Disc Ultegra",
        model == "Syapse Carbon Tiagra" ~ "Synapse Carbon Tiagra",
        model == "Supersix Evo Hi-Mod Utegra" ~ "Supersix Evo Hi-Mod Ultegra",
        TRUE ~ model
    ))

glimpse(bike_orderlines_tbl)
```



# Questions

# lubridate: Which month has the highest bike sales? (Difficulty = Medium)

- Start with `bike_orderlines_tbl`
- Select columns `order_date` and `total_price`
- Add a column called `month`
- Group by, summarize, and ungroup calculating the `sales`
- Arrange the sales values by month (Jan - Dec)
- Format the sales values as `dollar()`
- Adjust column names to title case

_What does this tell us about a time of year to focus marketing efforts?_

The highest sales are in April with over $8M.
```{r}
bike_orderlines_tbl %>% select(order_date, total_price) %>% 
    mutate(month = order_date %>% month(label = TRUE, abbr = FALSE)) %>% 
    group_by(month) %>% 
    summarise(sales = sum(total_price)) %>% 
    ungroup() %>%
    mutate(sales = scales::dollar(sales)) %>% 
    set_names(names(.) %>% str_to_title())
    


```


# stringr: What is the median orderline sales value by Bike Attribute? (Difficulty = Medium)

- Begin with `bike_orderlines`
- Select `model` and `total_price`
- Detect if string is present (e.g. "black inc")
- Groupby, summarize, and ungroup calculating the `median()` orderline
- Format numeric price as `dollar()` (Hint: investigate `largest_with_cents` argument)
- Rename column to evaluation string (e.g. "Black Inc") 


Evaluate "Black Inc". _What does this tell us about the "Black Inc" feature?_

Bike attribute "Black Inc" is an expensive feature. Bikes with this characteristics cost more and increase total sale.

```{r}
bike_orderlines_tbl %>% 
    select(model, total_price) %>% 
    
    #detecting the specific sentence in the model
    mutate(black_inc = model %>% str_to_lower() %>%  str_detect("black inc"),
           ultegra   = model %>% str_to_lower() %>% str_detect("ultegra"),
           disc      = model %>% str_to_lower() %>% str_detect("disc")) %>% 
    
    #made the calculation
    group_by(black_inc) %>% 
    summarise(median = median(total_price) %>% scales::dollar()) %>% 
    rename(`Black Inc`        = black_inc,
           `Median Orderline` = median)
    
    
```

Evaluate "Ultegra". _What does this tell us about the "Ultegra" feature?_

Bikes with "Ultegra" cost the same as other bikes. 

```{r}
bike_orderlines_tbl %>% 
    select(model, total_price) %>% 
    
    #detecting the specific sentence in the model
    mutate(black_inc = model %>% str_to_lower() %>%  str_detect("black inc"),
           ultegra   = model %>% str_to_lower() %>% str_detect("ultegra"),
           disc      = model %>% str_to_lower() %>% str_detect("disc")) %>% 
    
    #made the calculation
    group_by(ultegra) %>% 
    summarise(median = median(total_price) %>% scales::dollar()) %>% 
    rename(Ultegra            = ultegra,
           `Median Orderline` = median)
    
    
```


Evaluate "Disc" option. _What does this tell us about the "Disc" feature?_

The bikes that have "Disk" attribute usually cost more than similar bike without it. But this analysis - the mode of the total price between all bikes -  does not show it here. Comparison should be done between similar bikes that have and not have this attribute, but not with all bikes as other characteristics have more influence.

```{r}
bike_orderlines_tbl %>% 
    select(model, total_price) %>% 
    
    #detecting the specific sentence in the model
    mutate(black_inc = model %>% str_to_lower() %>%  str_detect("black inc"),
           ultegra   = model %>% str_to_lower() %>% str_detect("ultegra"),
           disc      = model %>% str_to_lower() %>% str_detect("disc")) %>% 
    
    #made the calculation
    group_by(disc) %>% 
    summarise(median = median(total_price) %>% scales::dollar()) %>% 
    rename(Disc               = disc,
           `Median Orderline` = median)

```


# stringr: What are the average, min, and max prices by Base Model? (Difficulty = High)

- Start with `bike_orderlines_tbl`
- Select distinct primary category, secondary category, model, and price (unit price, not total price)
- Create the base feature, `model_base` (Hint: Use the Feature Engineering code)
    - separate the models
    - Create a base feature that combines the appropriate parts (e.g. "Beast of the East")
- Remove any unnecessary columns (Hint: Deselect any columns matching `"model_[0-9]"`)
- Group by, summarize, and grouping (Hint: use `mean()`, `min()`, and `max()`)
- Arrange descending by average price
- Format any numeric columns as `dollar()` (Hint: Check out `largest_with_cents`) 
- Adjust the column names to title case

_What does this tell us about how bikes are priced?_
It shows the most expensive bikes in descending order. It also shows the range of prices for selected categories.
The most expensive bike is Scalpel-Si in Cross Country Race category.

```{r, warning=FALSE, message=FALSE}
bike_orderlines_tbl %>% distinct(category_1, category_2, model, price) %>% 
    
    #create a base feature 
    separate(
    col = model,
    into = str_c("model_", 1:7),
    sep = " ",
    remove = FALSE,
    fill = "right"   
    ) %>% 
    
    mutate(base_model = case_when(
        str_detect(model %>% str_to_lower(), "beast") ~ str_c(model_1, model_2, model_3, model_4, sep = " "),
        str_detect(model %>% str_to_lower(), "29|bad|fat|supersix") ~ str_c(model_1, model_2, sep = " "),
        TRUE ~ model_1
    )) %>% 
    select(-matches("model_[0-9]")) %>% 
    
#alternative just to select
   # select(category_1, category_2, model, base_model, price) 

    group_by(category_1, category_2, base_model) %>% 
    summarise(mean_price = mean(price),
              min_price  = min(price),
              max_price  = max(price)) %>% 
    ungroup() %>% 
    arrange(desc(mean_price)) %>% 
    mutate(mean_price  =  scales::dollar(mean_price, accuracy = 1),
           min_price   = scales::dollar(min_price),
           max_price   = scales::dollar(max_price)
    ) %>% 
    rename_with( ~ gsub("_", " ", .x, fixed = TRUE)) %>% 
    rename_with(~str_to_title(.) )
 
```



